Comparing summed probability distributions of shoreline and radiocarbon dates from the Stone Age on the Norwegian Skagerrak coast

### 2023-04-24 analysis/paper/paper.Rmd
ADMUR

```{r spduniform, echo = FALSE, fig.cap = "Monte Carlo simulation for shoreline and radiocarbon dates. A) Summed probability of shoreline dated sites (n = 934) compared to a uniform null model. B) Same as A, but with a exponential model. C) Summed probability of calibrated radiocarbon dates (n =, bins = 678) compared to a uniform null model. D) Same as C, but with a exponential null model.", out.width = "500px", fig.align="center", regfloat = TRUE}
knitr::include_graphics(here("analysis/figures/mc.png"))
```
Comparing the the directions and devitions from the three standard models, Figure appears to indicate some process of overall general increase or uniform distribution in the frequency of radiocarbon dates over time, contrasted to a general decrease in shoreline dated sites over time. However, with global p-values all below X,  all three models can be rejected as explaining the data they have been fit to, both for the SSPD and RSPD. Based on common demographic developments and previous suggestions, there is possibly some theoretical justification for expecting an exponential or logistic model to explain the shape of the RSPD, thus giving some further credence to interpreting deviations from this overall development as potentially meaningful signals of demographic developments. However, as introduced above, several authors have warned against over-interpreting such deviation. Furthermore, there are even less foundations on which to hypothesise what processes have dictated the development of site frequency over time, thus offering even further challenges to the interpretation of the SSPD and it's relationship with the null-models.

>>>>

```{r spduniform, echo = FALSE, fig.cap = "Monte Carlo simulation for shoreline and radiocarbon dates. A) Summed probability of shoreline dated sites (n = 934) compared to a uniform null model. B) Same as A, but with a exponential model. C) Summed probability of calibrated radiocarbon dates (n =, bins = 678) compared to a uniform null model. D) Same as C, but with a exponential null model.", out.width = "500px", fig.align="center", regfloat = TRUE}
knitr::include_graphics(here("analysis/figures/mc.png"))
```
Comparing the the directions and deviations from the three standard models, Figure appears to indicate some process of overall general increase or uniform distribution in the frequency of radiocarbon dates over time, contrasted to a general decrease in shoreline dated sites over time. However, with global p-values all below X,  all three models can be rejected as explaining the data they have been fit to, both for the SSPD and RSPD. Based on common demographic developments and previous suggestions, there is possibly some theoretical justification for expecting an exponential or logistic model to explain the shape of the RSPD, thus giving some further credence to interpreting deviations from this overall development as potentially meaningful signals of demographic developments. However, as introduced above, several authors have warned against over-interpreting such deviation. Furthermore, there are even less foundations on which to hypothesise what processes have dictated the development of site frequency over time, thus offering even further challenges to the interpretation of the SSPD and it's relationship with the null-models.

For the RSPD the rejected logistic model nonetheless offers the best approximation of the data. If taken as an indication of demographic developments this would indicate a first migration of people that is more substantial that what would be expected under the logistic null. Further deviations are then evident with the negative deviation and the poistive deviation c. These are, however, brief and can be left as little more than suggestions for possible meaningful demographic events.
